Learning Actionable Analytics (with applications for reducing defects and reducing runtimes)

### 2015-05-10 how.tex
Updates from ShareLaTeX

\section{Introduction}
Business  users   now demand   data mining tools
that  support a  business-level
interpretation of that data. For example,
at a  panel on software analytics at ICSE'12,
industrial practitioners lamented the state of the art in data mining
and software engineering~\cite{menzies12a}. Panelists commented that
``prediction is all well and good, but what about decision
making?''. That is, these panelists are more interested in the interpretations
and follow-up
that occurs {\em after} the mining, rather than just  the mining itself. For example:
\bi
\item Instead of just accepting  {\em predictions} on how many 
 software defects
to expect,  business users might now demand a {\em plan} to
reduce the likelihood of those defects;
\item Instead of just accepting {\em predictions} on the runtime
time of their software, business users might now demand
a {\em plan} to reduce that runtime.
\ei
In these two points, we say that a ``plan'' is a  set of proposed changes in order to better achieve some goal. 
Generating {\em plans} is a  very different task the standard
data mining tasks of {\em predicting} some discrete or numeric class.
Predictors
and classifiers only tell us what {\em is} while planners 
  tell us {\em what to change}. For the kind of business user
discussed above, planners are the preferred to mere prediction systems
since, as one gruff user once shouted at us:
\begin{quote}
{\em ``Don't tell me what \underline{is};
tell me what to \underline{change}.''}
\end{quote}
Generating plans is also differed to the recommender systems. 
discussed by Robillard and Walker~\cite{rob14}.  Recommendor systems
in software engineering
are less {\em directive} and more    {\em collaborative} than plan generators:
\bi
\item
Directive systems (such as those discussed in this paper) tell developers exactly what to do;
e.g. ``use these settings in a Makefile''. 
\item
On the other hand, the collaborative systems discussed by Robillard and Walker
are more suggestive guides that
help (e.g.) programmers focus on very small sets
with  very large libraries of code of documentation.
\ei
This paper presents HOW, a novel planner for learning actionable analytics.
In the experiments shown below, we see that HOW's plans can
significantly reduce the expected
value of:
\bi
\item The software defects found in  the object-oriented JAVA systems explored in 2010 by Jureczko et al.~\cite{jureczko10};
\item The runtime of the  software systems whose configurations
are explored in 2013 by Seigmund et al.~\cite{sven12};
\ei
In summary, this paper makes five contributions that are significantly different to
prior work:
\be
\item The HOW  instance-based planner of this paper has not been presented before.
\item The experiments of this paper show that HOW planner is useful for real-world tasks;
e.g. reducing the
expected values of the Jureczko software defects and the Seigmund software runtimes.
\item Apart from HOW, we  also offers a general framework for using  predictive system to build planners;
\item Further, we offer a  test to recognizes  when we should {\em not}  plan: specifically,
planning is
not recommended in domains with  poorly performing predictors;
\item Last, we  offer a general verification rig for assessing the value of plans generated within this rig.
\ee
The rest of this paper is structured as follows.
HOW's instance-based reasoning  makes its conclusions by
studying the space of examples. As discussed in \tion{mm},
this is a very different approach to
{\em model-based planners} that use the model
as sub-routine within a Monte Carlo simulation~\cite{me07f} or the evolutionary 
programming methods explored by the search-based software
engineering community~\cite{krall14,harman12dec}.

\section{HOW: Motivation and Methods}\label{sec:mm}

 \subsection{Motivation: Why ``HOW''?}
HOW was inspired by two     business issues raised about the  GALE optimizer~\cite{krall14}. 
A well-known effect in machine learning (and, in computer vision
and compression~\cite{levina04}) is that the $n$ dimensions of a data set can be represented by a  much
lower-dimensional approximation. GALE finds those $m \ll n$ dimensions then mutates examples
towards the better end of each dimension $m_i$. Since it explores $m \ll n$ dimensions,
GALE terminates much faster than standard optimiser (e.g. for one large model, 4 minutes versus 7 hours~\cite{krall14}).
Better yet, GALE's rapdily generated  solutions are competitive (or better) than those found by 
slower optimizers.

When GALE executes, it searches for better examples of data
 by (a)~inputting an initial population of data; then (b)~making
 extensive mutation to that data. In practice, GALE makes
100s to 100,000s of mutations. Each of these mutations
is reassessed using some domain model.
For certain users,this approach raises some issues.  One such user commented:
\begin{quote}
ISSUE~\#1: {\em ``What if I only want the minimum useful change, not
the maximum possible?''}
\end{quote}
Note here that this business user is asking for solutions  that do not require extensive 
changes to their current  approaches.

Later in that discussion, another business user asked:
\begin{quote}
ISSUE~\#2: {\em ``But what if I do not have a domain model of my business?''}
\end{quote}
This also is a legitimate issue.
Building and certifying a model  can  takes  long time, especially in software engineering. For  example, in previous work~\cite{me07f} we used models
developed by Boehm's group at the University of Southern California.
Some of those models took decades to develop and mature (from 1981~\cite{boehm81} to 2000~\cite{boehm00b}).

Also, even when there is an existing model, they can require
constant  maintenance lest they become out-dated. Elsewhere, we have described our
extensions to the USC models to enable reasoning about agile software developments. 
It took many months to implement and certify those extensions~\cite{me09i,me09j}.
The problem of model maintenance is another
motivation to look for alternate methods that can be automatically updated whenever new data arrives.

Further, sometimes  it is easier to access instances of a model's behaviour than the model
itself. For example, in prior work with Martin  Feather from the Jet Propulsion 
Laboratory~\cite{fea02a},  our research partner could not share a
propriety model from within the NASA firewalls. However, they could share 
logs of the input/output of that model.

These two issues are not just problems for GALE, but also for every
other model-based optimization algorithm  that makes
extensive use of mutation  and domain models such as
NSGA-II, NSGA-III, SPEA2, IBEA, PSO, DE, MOEA/D, etc.~\cite{deb00a,zit02,zit04,%
deb14,Cui2005a,storn97,zhang07:TEC}. Accordingly, this paper seeks
an optimization methods that make minimal changes to instances,
while avoiding the problem of models that are missing, slow
to build, hard to maintain, or difficult to access.

\subsection{Methods: How to Build ``HOW''?}

A general model for {\em predictors} is as follows. 
During testing, each test instance gains a prediction 
by finding which region of the data it falls into.
During training, those regions are inferred from
divisions that ``chunk'' together regions with instances
with similar properties. These dividers can be categorical (as in the case of decision trees) or probabilistic (as in the
case of Naive Bayes) or learned via some distance minimization
process 
(as in the case of support vector machines) or generated
directly by some clustering algorithm.

@@

An important design decision with HOW 
is how to determine the exact {\em direction} and {\em magnitude} of the
displacement.  Following on from the discussion around   ISSUE~\#1,
we note that a design goal of HOW is
{\em minimal displacements} of project data.
To implement minimal displacement:
\bi
\item
HOW sorts  the independent variables
using the procedure of  \fig{where}d. It then only
displaces the    best $B$\%  of that sort, while leaving the rest untouched
(see \fig{howcode}, line 38).
In  the experiments shown below,  we explore   $B \in \{25,50,100\}\%$.
\item
HOW also  limits displacement
to very small regions in the space of the training data (never more that the separation
of the two closest clusters).  HOW's clusterer divides  data into many   small clusters containing
$\sqrt{N}$ of the population (using the {\tt cluster} function, \fig{howcode}, line 76).
HOW   limits displacement to  \mbox{$0 \le F \le c$}, where
$c$ is the distance between the two closest clusters. 
In  the experiments shown below,  we explore
\mbox{$F\in \{0.25, 0.5, 0.75\}$}.
\ei
Turning now to the displacement {\em direction},
our experiments show that it useful to  carefully reflect on exactly where HOW should  push each test instance.
HOW replaces each cluster   with one {\em exemplar} that best
characterizes the data in that cluster. It then displaces the
test data towards that exemplar, as shown in \fig{howcode}, line 19. In summary, 
If  $Z_j$ is an independent variable within the
test instance $Z$ and  $X$ is the exemplar found at training time,
then we displace $Z$ as follows:
\[Z_j =  Z_j + F*(X_j - Z_j)\]
Here, $F$ is the  magnitude controller, discussed above.
The experiments of this work showed that different kinds of data require
different kinds of exemplars.
\bi
\item
For the Seigmund data,
the objective score is the runtime of the test suite of a compiled program.
For this kind of data with numeric objectives,
the exemplar  is the instance in a cluster
with  best objective score.
\item
For the Jureczko data, the objective score
is a boolean that is true if defects have been noted
for that class. 
For this kind of data with discrete objectives,
the data, 
has less differentiation between the objective scores
so we set the exemplar to  
an instance synthesized from the median values of the instances
in that cluster.
\ei

@@

\subsection{How to Assess ``HOW''}\label{sect:assess}
In the experiments shown below,  HOW comments  on how to reduce
software defects by adjusting certain static code parameters of the code\footnote{Such parameters can be manually adjusted by programmers or (for large scale changes) automatically adjusted using code refactoring
tools, as shown by Binkely et al. in 2006~\cite{Binkley2006} and more recently by Mkaouer et al. in 2015~\cite{Mkaouer15}.}. How are we to verify that HOW's recommendations actually work?

@@

\section{When Not to Plan: the Problem of Poor Predictors}

@@

\section{Results}

@@

\begin{figure*}[htbp!]

\noindent\begin{minipage}{0.5\textwidth}
  {\scriptsize \textbf{Ant}\\}
  {\scriptsize  \begin{tabular}{{l@{~~~~}l@{}r@{~~~~}r@{~~~~}c}}
      \arrayrulecolor{darkgray}
      \rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
  1 & 0.75 &    -0.63  &  0.08 & \quart{2}{7}{5}{144} \\
  1 & 0.75, weighting &    -0.6  &  0.2 & \quart{0}{16}{8}{144} \\
\hline  2 & 0.5, weighting &    -0.42  &  0.11 & \quart{17}{10}{23}{144} \\
  2 & 0.5 &    -0.42  &  0.11 & \quart{19}{9}{23}{144} \\
  2 & 0.75, weighting, Prune=50\% &    -0.4  &  0.11 & \quart{20}{9}{25}{144} \\
\hline  3 & 0.25 &    -0.28  &  0.09 & \quart{31}{7}{35}{144} \\
  3 & 0.25, weighting &    -0.28  &  0.08 & \quart{31}{7}{35}{144} \\
\hline  4 & 0.5, weighting, Prune=50\% &    -0.25  &  0.06 & \quart{35}{5}{38}{144} \\
\hline  5 & 0.25, weighting, Prune=50\% &    -0.19  &  0.05 & \quart{41}{4}{43}{144} \\
\hline  6 & Baseline &    -0.11  &  0.03 & \quart{47}{2}{49}{144} \\
\hline \end{tabular}}
\end{minipage}
\noindent\begin{minipage}{0.5\textwidth}
  \flushleft
  {\scriptsize \textbf{Camel}\\}
{\scriptsize  \begin{tabular}{{l@{~~~~}l@{}r@{~~~~}r@{~~~~}c}}
\arrayrulecolor{darkgray}
\rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
  1 &  0.75, weighting, Prune=50\% &    -0.41  &  0.1 & \quart{2}{14}{9}{205} \\
  1 &  0.75 &    -0.41  &  0.23 & \quart{1}{32}{9}{205} \\
  1 &  0.75, weighting &    -0.39  &  0.16 & \quart{0}{22}{12}{205} \\
\hline  2 &  0.5, weighting, Prune=50\% &    -0.27  &  0.09 & \quart{19}{12}{29}{205} \\
  2 &  0.5 &    -0.26  &  0.12 & \quart{20}{17}{30}{205} \\
  2 &  0.5, weighting &    -0.24  &  0.09 & \quart{24}{13}{33}{205} \\
\hline  3 &  0.25, weighting, Prune=50\% &    -0.2  &  0.05 & \quart{36}{7}{38}{205} \\
  3 &  0.25, weighting &    -0.19  &  0.07 & \quart{36}{9}{40}{205} \\
  3 &  0.25 &    -0.17  &  0.06 & \quart{37}{8}{43}{205} \\
\hline  4 &  Baseline &    -0.15  &  0.06 & \quart{41}{8}{45}{205} \\
\hline \end{tabular}}
\end{minipage}

\noindent\begin{minipage}{0.5\textwidth}
  \flushleft
  {\scriptsize \textbf{Ivy}\\[-0.05cm]}
{\scriptsize  \begin{tabular}{{l@{~~~~}l@{}r@{~~~~}r@{~~~~}c}}
\arrayrulecolor{darkgray}
\rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
  1 &     0.75 &    -0.57  &  0.17 & \quart{0}{16}{6}{163} \\
  1 &   0.75, weighting &    -0.55  &  0.11 & \quart{3}{11}{8}{163} \\
\hline  2 &   0.75, weighting, Prune=50\% &    -0.38  &  0.11 & \quart{19}{11}{25}{163} \\
  2 &    0.5, weighting &    -0.36  &  0.12 & \quart{23}{12}{27}{163} \\
  2 &   0.5 &    -0.33  &  0.11 & \quart{25}{11}{30}{163} \\
\hline  3 &   0.5, weighting, Prune=50\% &    -0.27  &  0.08 & \quart{30}{8}{36}{163} \\
\hline  4 &   0.25, weighting &    -0.2  &  0.07 & \quart{41}{7}{43}{163} \\
  4 &     0.25 &    -0.16  &  0.04 & \quart{43}{4}{47}{163} \\
  4 &     Baseline &    -0.16  &  0.0 & \quart{47}{0}{47}{163} \\
\hline  5 &   0.25, weighting, Prune=50\% &    -0.15  &  0.04 & \quart{45}{4}{48}{163} \\
\hline \end{tabular}}
\end{minipage}
\noindent\begin{minipage}{0.5\textwidth}
  \flushleft
  {\scriptsize \textbf{Jedit}\\}
  {\scriptsize  \begin{tabular}{{l@{~~~~}l@{}r@{~~~~}r@{~~~~}c}}
\arrayrulecolor{darkgray}
\rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
  1 &  0.75, weighting, Prune=50\% &    -0.41  &  0.18 & \quart{0}{20}{8}{172} \\
  1 &  0.75 &    -0.35  &  0.26 & \quart{0}{30}{15}{172} \\
\hline  2 &  0.75, weighting &    -0.25  &  0.17 & \quart{16}{20}{26}{172} \\
  2 &  0.5, weighting, Prune=50\% &    -0.21  &  0.1 & \quart{23}{11}{31}{172} \\
\hline  3 &   0.5 &    -0.17  &  0.08 & \quart{32}{9}{36}{172} \\
  3 &  0.5, weighting &    -0.14  &  0.09 & \quart{31}{10}{39}{172} \\
\hline  4 &  0.25, weighting, Prune=50\% &    -0.12  &  0.06 & \quart{39}{7}{41}{172} \\
\hline  5 &  0.25 &    -0.08  &  0.05 & \quart{43}{5}{46}{172} \\
  5 &  0.25, weighting &    -0.08  &  0.06 & \quart{43}{6}{46}{172} \\
\hline  6 &  Baseline &    -0.06  &  0.01 & \quart{47}{1}{48}{172} \\
\hline \end{tabular}}
\end{minipage}

\noindent\begin{minipage}{0.5\textwidth}
  \flushleft
  {\scriptsize \textbf{Log4j}\\}
  {\scriptsize  \begin{tabular}{{l@{~~~~}l@{}r@{~~~~}r@{~~~~}c}}
\arrayrulecolor{darkgray}
\rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
  1 &  0.75, weighting, Prune=50\% &    -0.19  &  0.13 & \quart{0}{28}{15}{273} \\
\hline  2 &  0.75 &    -0.13  &  0.06 & \quart{23}{13}{28}{273} \\
  2 &  0.5, weighting, Prune=50\% &    -0.13  &  0.05 & \quart{23}{11}{28}{273} \\
  2 &  0.75, weighting &    -0.12  &  0.1 & \quart{15}{21}{30}{273} \\
\hline  3 &  0.5, weighting &    -0.1  &  0.03 & \quart{28}{6}{34}{273} \\
\hline  4 &   0.5 &    -0.1  &  0.07 & \quart{28}{15}{34}{273} \\
  4 &  0.25 &    -0.09  &  0.04 & \quart{34}{9}{36}{273} \\
  4 &  0.25, weighting, Prune=50\% &    -0.09  &  0.04 & \quart{34}{9}{36}{273} \\
  4 &  0.25, weighting &    -0.06  &  0.04 & \quart{34}{9}{43}{273} \\
\hline  5 &  Baseline &    -0.06  &  0.03 & \quart{43}{6}{43}{273} \\
\hline \end{tabular}}
\end{minipage}
\noindent\begin{minipage}{0.5\textwidth}
  \flushleft
  {\scriptsize \textbf{Lucene}\\}
  {\scriptsize  \begin{tabular}{{l@{~~~~}l@{}r@{~~~~}r@{~~~~}c}}
\arrayrulecolor{darkgray}
\rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
  1 &  0.75, weighting &    -0.15  &  0.09 & \quart{0}{29}{9}{393} \\
  1 &  0.75 &    -0.13  &  0.09 & \quart{0}{29}{16}{393} \\
  1 &  0.75, weighting, Prune=50\% &    -0.12  &  0.08 & \quart{6}{27}{19}{393} \\
\hline  2 &  Baseline &    -0.1  &  0.03 & \quart{26}{10}{26}{393} \\
  2 &  0.5, weighting, Prune=50\% &    -0.1  &  0.03 & \quart{23}{10}{26}{393} \\
\hline  3 &  0.5, weighting &    -0.08  &  0.06 & \quart{23}{20}{33}{393} \\
  3 &  0.25, weighting, Prune=50\% &    -0.07  &  0.03 & \quart{33}{10}{36}{393} \\
  3 &  0.5 &    -0.07  &  0.05 & \quart{29}{17}{36}{393} \\
\hline  4 &  0.25, weighting &    -0.04  &  0.04 & \quart{36}{13}{46}{393} \\
  4 &  0.25 &    -0.04  &  0.02 & \quart{39}{7}{46}{393} \\
\hline \end{tabular}}
\end{minipage}

>>>>

\section{Introduction}
Business  users   now demand   data mining tools
that  support a  business-level
interpretation of that data. For example,
at a  panel on software analytics at ICSE'12,
industrial practitioners lamented the state of the art in data mining
and software engineering~\cite{menzies12a}. Panelists commented that
``prediction is all well and good, but what about decision
making?''. That is, these panelists are more interested in the interpretations
and follow-up
that occurs {\em after} the mining, rather than just  the mining itself. For example:
\bi
\item 
Instead of just accepting  {\em predictions} on how many 
 software defects
to expect,  business users might now demand a {\em plan} to
reduce the likelihood of those defects;
\item Instead of just accepting {\em predictions} on the runtime
time of their software, business users might now demand
a {\em plan} to reduce that runtime.
\ei
In this paper, we say   a ``plan'' is a  set of changes
proposed  in order to better achieve some goal. 
Specifically, for every $Z$  that requests a plan $\delta$, we want
to adjust the attributes $Z_j$ as follows:
\begin{equation}\label{eq:one}
Z_j = Z_j + \delta_j
\end{equation}
The usual way such plans are generated in the SE literature
is via some search-based method. Recently, we have made  some progress
in this area with the  GALE optimizer~\cite{krall14}. 
Like other search-based algorithms,
 GALE executes by proposing multiple $\delta$s
 (one for each generation in an evolutionary program).
 In practice, GALE makes
 extensive changes to initial sample of project data
 (perhaps 
100s to 100,000s of mutations). Each of these mutations
is reassessed using some domain model.
GALE runs faster than standard search-based SE tools~\cite{deb00a,zit02} since
it explores  the principle components
 of the project data (rather than   the raw data). Since it explores
 a smaller space, it also
 terminates much faster than standard optimiser; e.g. for one large model, 4 minutes versus 7 hours~\cite{krall14}.

When presented to some  users, the GALE approach was challenged by
two business users.   Note that the following issues
challenge not just GALE, but also  any  optimization algorithm  that makes
extensive use of mutation  and domain models such as
NSGA-II, NSGA-III, SPEA2, IBEA, particle swarm optimization, MOEA/D, etc.~\cite{deb00a,zit02,zit04,%
deb14,Cui2005a,zhang07:TEC}.

One  user asked:
\begin{quote}
ISSUE~\#1: {\em ``What if I only want the minimum useful change, not
the maximum possible?''}
\end{quote}
Note here that this business user is asking for solutions  that do not require extensive 
changes to their current  approaches.

Later in that discussion, another business user asked:
\begin{quote}
ISSUE~\#2: {\em ``But what if I do not have a domain model of my business?''}
\end{quote}
This also is a legitimate issue.
Building and certifying a model  can  takes  long time, especially in software engineering. For  example, in previous work~\cite{me07f} we used models
developed by Boehm's group at the University of Southern California.
Some of those models took decades to develop and mature (from 1981~\cite{boehm81} to 2000~\cite{boehm00b}). 
Also, even when there is an existing model, they can require
constant  maintenance lest they become out-dated. Elsewhere, we have described our
extensions to the USC models to enable reasoning about agile software developments. 
It took many months to implement and certify those extensions~\cite{me09i,me09j}.
The problem of model maintenance is another
motivation to look for alternate methods that can be automatically updated whenever new data arrives.
Further, sometimes  it is easier to access instances of a model's behaviour than the model
itself. For example, in prior work with Martin  Feather from the Jet Propulsion 
Laboratory~\cite{fea02a},  our research partner could not share a
propriety model from within the NASA firewalls. However, they could share 
logs of the input/output of that model.

Every challenge is an opportunity to explore  a new approach.
If we  restrict $\delta$ to very small changes, then any 
predictive data miner applied to the original data could also be used
to assessed the mutated data (proviso: we have to guarantee that
the mutations do not leave the space of examples present in the original data).
Note that this approach addresses ISSUE~\#1 since it only uses minimal changes.
It also addresses  ISSUE~\#2 since, when lack a domain model, we automatically
create one (via data mining).

This paper presents HOW, a novel planner for learning actionable analytics via
minimal changes. 
In the experiments shown below, we see that HOW's plans  can significantly
reduce the  expected
value of:
\bi
\item The software defects found in  the object-oriented JAVA systems explored in 2010 by Jureczko et al.~\cite{jureczko10};
\item The runtime of the  software systems whose configurations
are explored in 2013 by Seigmund et al.~\cite{sven12};
\ei
This paper makes five contributions:
\be
\item The HOW  instance-based planner (which has not been presented before).
\item The experiments of this paper show that HOW planner is useful for real-world tasks;
e.g. reducing the
expected values of the Jureczko software defects and the Seigmund software runtimes.
\item Apart from HOW, we  also offers a general framework for using  predictive system to build planners;
\item Further, we offer a  test to recognizes  when we should {\em not}  plan: specifically,
planning is
not recommended in domains with  poorly performing predictors;
\item Last, we  offer a general verification rig for assessing the value of plans generated within this rig.
\ee

\section{How to Build ``HOW''?}\label{sec:mm}

\subsection{Design}
A general model for {\em predictors} is as follows. 
During testing, each test instance gains a prediction 
by finding which region of the data it falls into.
During training, those regions are inferred from
divisions that ``chunk'' together regions with instances
with similar properties. These dividers can be categorical (as in the case of decision trees) or probabilistic (as in the
case of Naive Bayes) or learned via some distance minimization
process 
(as in the case of support vector machines) or generated
directly by some clustering algorithm.

@@

An important design decision with HOW 
is how to determine the exact {\em direction} and {\em magnitude} of the
displacement.  Following on from the discussion around   ISSUE~\#1,
we note that a design goal of HOW is
{\em minimal displacements} of project data.
To implement minimal displacement:
\bi
\item
HOW sorts  the independent variables
using the procedure of  \fig{where}d. It then only
displaces the    best $B$\%  of that sort, while leaving the rest untouched
(see \fig{howcode}, line 38).
In  the experiments shown below,  we explore   $B \in \{25,50,100\}\%$.
\item
HOW also  limits displacement
to very small regions in the space of the training data (never more that the separation
of the two closest clusters).  HOW's clusterer divides  data into many   small clusters containing
$\sqrt{N}$ of the population (using the {\tt cluster} function, \fig{howcode}, line 76).
HOW   limits displacement to  \mbox{$0 \le F \le c$}, where
$c$ is the distance between the two closest clusters. 
In  the experiments shown below,  we explore
\mbox{$F\in \{0.25, 0.5, 0.75\}$}.
\ei
Turning now to the displacement {\em direction},
our experiments show that it useful to  carefully reflect on exactly where HOW should  push each test instance.
HOW replaces each cluster   with one {\em exemplar} that best
characterizes the data in that cluster. It then displaces the
test data towards that exemplar, as shown in \fig{howcode}, line 19. In summary, 
If  $Z_j$ is an independent variable within the
test instance $Z$ and  $X$ is the exemplar found at training time,
then we displace $Z$ as follows:
\[Z_j =  Z_j + F*(X_j - Z_j)\]
Here, $F$ is the  magnitude controller, discussed above. Note that this
this equations satisfies our definition of a ``plan'' as offered in
\eq{one} in the {\em Introduction}

The experiments of this work showed that different kinds of data require
different kinds of exemplars.
\bi
\item
For the Jureczko data, the objective score
is a boolean that is true if defects have been noted
for that class. 
For this kind of data with discrete objectives,
the data, 
has very little to  differentiate between the objective scores
so we just set the exemplar to  
an instance synthesized from the median values of the instances
in that cluster.
\item
For the Seigmund data,
the objective score is the runtime of the test suite of a compiled program.
For this kind of data with numeric objectives,
the exemplar  is the instance in a cluster
with  best objective score.
\ei

@@

\subsection{Assessment}\label{sect:assess}
In the experiments shown below,  HOW comments  on how to reduce
software defects by adjusting certain static code parameters of the code\footnote{Such parameters can be manually adjusted by programmers or (for large scale changes) automatically adjusted using code refactoring
tools, as shown by Binkely et al. in 2006~\cite{Binkley2006} and more recently by Mkaouer et al. in 2015~\cite{Mkaouer15}.}. How are we to verify that HOW's recommendations actually work?

@@

\subsection{When Not to Plan: the Problem of Poor Predictors}

@@

\section{Experiments}

@@

\begin{figure}[htbp!]

\noindent\begin{minipage}{0.5\textwidth}
  {\scriptsize \textbf{Ant}\\[0.1cm]}
  {\scriptsize  \begin{tabular}{{l@{~~~~}l@{}r@{~~~~}r@{~~~~~}c@{}r}}
      \arrayrulecolor{darkgray}
      \rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} && \textbf{Bugs}\\
  1 &  0.75-weight &    0.71  &  0.19 & \quart{0}{25}{10}{49} & 62\\
  1 &         0.75 &    0.72  &  0.14 & \quart{6}{19}{12}{49} & 64\\
\hline  2 &          0.5 &    0.86  &  0.08 & \quart{25}{11}{31}{49} & 77 \\
  2 &   0.5-weight &    0.88  &  0.14 & \quart{20}{19}{33}{49} & 77\\
\hline  3 & 0.5-weight-B=50\% &    0.91  &  0.08 & \quart{32}{11}{37}{49} & 80\\
  3 & 0.75-weight-B=50\% &    0.92  &  0.09 & \quart{31}{12}{39}{49} & 82\\
  3 &  0.25-weight &    0.92  &  0.04 & \quart{36}{5}{39}{49} & 81\\
\hline  4 &         0.25 &    0.93  &  0.05 & \quart{37}{7}{40}{49} & 83\\
  4 & 0.75-weight-B=25\% &    0.94  &  0.09 & \quart{36}{12}{41}{49} & 82\\
\hline  5 & 0.25-weight-B=50\% &    0.97  &  0.03 & \quart{44}{4}{45}{49} & 86\\
  5 & 0.5-weight-B=25\% &    0.98  &  0.06 & \quart{40}{8}{47}{49} & 86 \\
\hline  6 & 0.25-weight-B=25\% &    0.99  &  0.01 & \quart{48}{1}{48}{49} & 87\\
\hline  7 &         Base &    1.0  &  0.0 & \quart{49}{0}{49}{49} & 90\\
\hline \end{tabular}}
\end{minipage}\\[0.1cm]

\noindent\begin{minipage}{0.5\textwidth}
  \flushleft
  {\scriptsize \textbf{Ivy}\\[0.1cm]}
{\scriptsize  \begin{tabular}{{l@{~~~~}l@{}r@{~~~~}r@{~~~~}c@{}r}}
\arrayrulecolor{darkgray}
\rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & & \textbf{Bugs}\\
  1 &  0.75-weight &    0.75  &  0.18 & \quart{0}{29}{8}{49} & 50\\
  1 &         0.75 &    0.79  &  0.12 & \quart{8}{20}{14}{49} & 53\\
\hline  2 &          0.5 &    0.84  &  0.1 & \quart{18}{16}{23}{49} & 56\\
  2 &   0.5-weight &    0.84  &  0.08 & \quart{14}{14}{23}{49} & 56\\
  2 & 0.75-weight-B=50\% &    0.84  &  0.09 & \quart{18}{15}{23}{49} & 56 \\
\hline  3 & 0.5-weight-B=50\% &    0.91  &  0.05 & \quart{29}{9}{34}{49} & 61\\
\hline  4 & 0.25-weight-B=50\% &    0.96  &  0.01 & \quart{43}{1}{43}{49} & 64\\
  4 & 0.75-weight-B=25\% &    0.96  &  0.05 & \quart{39}{9}{43}{49} & 65\\
  4 &         0.25 &    0.97  &  0.01 & \quart{43}{1}{44}{49} & 65\\
  4 &  0.25-weight &    0.97  &  0.05 & \quart{39}{9}{44}{49} & 65 \\
  4 & 0.5-weight-B=25\% &    0.97  &  0.03 & \quart{43}{5}{44}{49} & 65 \\
\hline  5 & 0.25-weight-B=25\% &    0.99  &  0.01 & \quart{48}{1}{48}{49} & 66\\
\hline  
6 &         Base &    1.0  &  0.0 & \quart{49}{0}{49}{49} & 67\\
\hline \end{tabular}}
\end{minipage}\\[0.1cm]

\noindent\begin{minipage}{0.5\textwidth}
  \flushleft
  {\scriptsize \textbf{Jedit}\\[0.1cm]}
  {\scriptsize  \begin{tabular}{{l@{~~~~}l@{}r@{~~~~}r@{~~~~}c@{}r}}
\arrayrulecolor{darkgray}
\rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & & \textbf{Bugs}\\\hline
  1 &         0.75 &    0.64  &  0.28 & \quart{0}{27}{14}{49} & 110\\
\hline  2 &  0.75-weight &    0.76  &  0.21 & \quart{13}{21}{26}{49}& 131 \\
  2 &   0.5-weight &    0.76  &  0.16 & \quart{20}{16}{26}{49} & 130\\
  2 &          0.5 &    0.79  &  0.17 & \quart{19}{17}{29}{49} & 137\\
\hline  3 & 0.75-weight-B=50\% &    0.85  &  0.12 & \quart{30}{12}{35}{49} & 146\\
  3 &  0.25-weight &    0.87  &  0.07 & \quart{34}{7}{37}{49} & 152\\
\hline  4 &         0.25 &    0.93  &  0.09 & \quart{37}{9}{43}{49} & 161\\
  4 & 0.5-weight-B=50\% &    0.93  &  0.05 & \quart{41}{5}{43}{49} & 162\\
\hline  5 & 0.25-weight-B=50\% &    0.97  &  0.03 & \quart{45}{3}{47}{49} & 167\\
  5 & 0.75-weight-B=25\% &    0.97  &  0.02 & \quart{46}{2}{47}{49} & 168\\
\hline  6 & 0.5-weight-B=25\% &    0.98  &  0.02 & \quart{47}{2}{48}{49} & 170\\
  6 & 0.25-weight-B=25\% &    0.98  &  0.01 & \quart{48}{1}{48}{49} & 171\\
\hline  7 &         Base &    1.0  &  0.0 & \quart{49}{0}{49}{49} & 175\\
\hline \end{tabular}}
\end{minipage}\\[0.1cm]

\noindent\begin{minipage}{0.5\textwidth}
  \flushleft
  {\scriptsize \textbf{Lucene}\\[0.1cm]}
  {\scriptsize  \begin{tabular}{{l@{~~~~}l@{}r@{~~~~}r@{~~~~}c@{}r}}
\arrayrulecolor{darkgray}
\rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & & \textbf{Bugs}\\
  1 &  0.75-weight &    0.66  &  0.2 & \quart{0}{21}{13}{49} & 97 \\
\hline  2 & 0.75-weight-B=50\% &    0.7  &  0.1 & \quart{14}{10}{17}{49} & 103 \\
\hline  3 &         0.75 &    0.77  &  0.3 & \quart{9}{33}{24}{49} & 113\\
  3 & 0.75-weight-B=25\% &    0.8  &  0.09 & \quart{21}{10}{28}{49} & 117\\
  3 &   0.5-weight &    0.81  &  0.09 & \quart{26}{9}{29}{49} & 119\\
  3 &          0.5 &    0.84  &  0.1 & \quart{26}{10}{32}{49} & 123\\
\hline  4 & 0.5-weight-B=50\% &    0.88  &  0.07 & \quart{34}{8}{36}{49} & 128\\
  4 & 0.25-weight-B=25\% &    0.89  &  0.07 & \quart{36}{8}{38}{49} & 130 \\
  4 &  0.25-weight &    0.92  &  0.04 & \quart{39}{4}{41}{49} & 133 \\
  4 &         0.25 &    0.93  &  0.07 & \quart{36}{8}{42}{49} & 136 \\
\hline  5 & 0.25-weight-B=50\% &    0.95  &  0.04 & \quart{42}{4}{44}{49} & 139 \\
  5 & 0.5-weight-B=25\% &    0.96  &  0.03 & \quart{43}{3}{45}{49} &141 \\
\hline  6 &         Base &    1.0  &  0.0 & \quart{49}{0}{49}{49} & 146\\
\hline \end{tabular}}
\end{minipage}\\[0.1cm]
